Hamby Data and `rtrees`

### 2020-09-07 journals/01-hamby_data_and_models/01-hamby_data_and_models.Rmd
updates for preparing paper for submission

```{r}
# Load libraries
library(bulletr)
library(cowplot)
library(ggConvexHull) #devtools::install_github("cmartin/ggConvexHull")
library(GGally)
library(ggpcp)
library(glmnet)
library(plotly)
library(randomForest)
library(tidyverse)
```

@@

Both the training and the testing data will require the use of a vector of the names of the features used in the `rtrees` random forest included in the bulletr library. I created such a vector to use in this journal, and this vector is printed below. It includes nine features.

```{r}
# Obtain features used when fitting the rtrees random forest
rf_features <- rownames(bulletr::rtrees$importance)
rf_features
```

@@

# Cleaning the testing data
hamby173and252_train_cleaning <- hamby173and252_raw %>%
  mutate(study1 = fct_recode(study1, "Hamby173" = "Hamby44"),
         study2 = fct_recode(study2, "Hamby173" = "Hamby44"),
         bullet1 = factor(ifelse(barrel1 %in% letters,
                                 as.character(barrel1),
                                 as.character(bullet1))),
         barrel1 = factor(ifelse(barrel1 %in% letters, 
                                 as.character("Unknown"), 
                                 as.character(barrel1))),
         bullet2 = factor(ifelse(barrel2 %in% letters,
                                 as.character(barrel2),
                                 as.character(bullet2))),
         barrel2 = factor(ifelse(barrel2 %in% letters, 
                                 as.character("Unknown"), 
                                 as.character(barrel2))),
         land1 = factor(land1),
         land2 = factor(land2),
         rfscore = predict(bulletr::rtrees, hamby173and252_raw %>%
                             select(rf_features), 
                           type = "prob")[,2]) %>%
  rename(samesource = match) %>%
  select(study1, barrel1, bullet1, land1, study2, barrel2, bullet2, land2,
         rf_features, samesource, rfscore)
```

@@

```{r}
# Compare the dimensions of my training data and the number of predictions from rtrees
dim(hamby173and252_train_cleaning)
dim(predict(bulletr::rtrees, type = "prob"))
```

@@

```{r}
# Compare the dimensions of my further cleaned training data and the
# number of predictions from rtrees
dim(hamby173and252_train)
dim(predict(bulletr::rtrees, type = "prob"))
```

@@

```{r eval = FALSE}
# Save the datasets and response variables as .csv files
write.csv(hamby173and252_train, "../../data/hamby173and252_train.csv", row.names = FALSE)
```

@@

# Join the two cleaned Hamby 224 sets into one testing set
hamby224_test <- suppressWarnings(bind_rows(hamby224_set1_cleaned,
                                            hamby224_set11_cleaned)) %>%
  mutate(set = factor(set),
         bullet1 = factor(bullet1),
         bullet2 = factor(bullet2),
         land1 = factor(land1),
         land2 = factor(land2)) %>%
  right_join(combinations, by = c("set", "land1", "land2", "bullet1", "bullet2")) %>%
  filter(!(bullet1 == "Questioned" & bullet2 == "Known 1"),
         !(bullet1 == "Questioned" & bullet2 == "Known 2"),
         !(bullet1 == "Known 2" & bullet2 == "Known 1")) %>%
  arrange(rfscore) %>%
  mutate(case = factor(1:length(study))) %>%
  select(case, study:samesource)
```

@@

```{r eval = FALSE}
# Save the test data as a .csv file
write.csv(hamby224_test, "../../data/hamby224_test.csv", row.names = FALSE)
```

@@

The random forest model that was fit to the Hamby training dataset is discussed in the paper "Automatic Matching of Bullet Land Impressions" by E. Hare Et. Al. The model is available through the GitHub version of the `bulletr` package. The link to the GitHub repository for `bulletr` is https://github.com/CSAFE-ISU/bulletr.

The random forest model is stored in the package `bulletr` as a data object called `rtrees`. The paper does not include many details about the model fitting process, so I looked at some of the model features, which are printed below. These values told me that the model was fit using the predictive features of `ccf`, `rough_cor`, `D`, `sd_D`, `matches`, `mismatches`, `cms`, `non_cms`, and `sum_peaks` with 300 trees and the parameter of `mtry` set to 3, which is the "number of variables randomly sampled as candidates at each split" as explained in the `randomforest` package documentation.

@@

```{r}
# Create a dataframe with the random forest features ordered by importance
rf_features_ordered <- data.frame(feature = rownames(bulletr::rtrees$importance), 
                                  MeanDecreaseGini = bulletr::rtrees$importance) %>%
  arrange(desc(MeanDecreaseGini))
```

>>>>

```{r}
# Load libraries
library(bulletxtrctr)
library(cowplot)
library(ggConvexHull) #devtools::install_github("cmartin/ggConvexHull")
library(GGally)
library(ggpcp)
library(glmnet)
library(plotly)
library(randomForest)
library(tidyverse)
```

@@

Both the training and the testing data will require the use of a vector of the names of the features used in the `rtrees` random forest included in the `bulletxtrctr` library. I created such a vector to use in this journal, and this vector is printed below. It includes nine features.

```{r}
# Obtain features used when fitting the rtrees random forest
rf_features <- rownames(bulletxtrctr::rtrees$importance)
rf_features
```

@@

# Cleaning the testing data
hamby173and252_train_cleaning <- hamby173and252_raw %>%
  mutate(study1 = fct_recode(study1, "Hamby173" = "Hamby44"),
         study2 = fct_recode(study2, "Hamby173" = "Hamby44"),
         bullet1 = factor(ifelse(barrel1 %in% letters,
                                 as.character(barrel1),
                                 as.character(bullet1))),
         barrel1 = factor(ifelse(barrel1 %in% letters, 
                                 as.character("Unknown"), 
                                 as.character(barrel1))),
         bullet2 = factor(ifelse(barrel2 %in% letters,
                                 as.character(barrel2),
                                 as.character(bullet2))),
         barrel2 = factor(ifelse(barrel2 %in% letters, 
                                 as.character("Unknown"), 
                                 as.character(barrel2))),
         land1 = factor(land1),
         land2 = factor(land2),
         rfscore = predict(bulletxtrctr::rtrees, hamby173and252_raw %>%
                             select(rf_features), 
                           type = "prob")[,2]) %>%
  rename(samesource = match) %>%
  select(study1, barrel1, bullet1, land1, study2, barrel2, bullet2, land2,
         rf_features, samesource, rfscore)
```

@@

```{r}
# Compare the dimensions of my training data and the number of predictions from rtrees
dim(hamby173and252_train_cleaning)
dim(predict(bulletxtrctr::rtrees, type = "prob"))
```

@@

```{r}
# Compare the dimensions of my further cleaned training data and the
# number of predictions from rtrees
dim(hamby173and252_train)
dim(predict(bulletxtrctr::rtrees, type = "prob"))
```

@@

```{r}
# Save the datasets and response variables as .csv files
write.csv(hamby173and252_train, "../../../data/hamby173and252_train.csv", row.names = FALSE)
```

@@

# Join the two cleaned Hamby 224 sets into one testing set
hamby224_test <- suppressWarnings(bind_rows(hamby224_set1_cleaned,
                                            hamby224_set11_cleaned)) %>%
  mutate(set = factor(set),
         bullet1 = factor(bullet1),
         bullet2 = factor(bullet2),
         land1 = factor(land1),
         land2 = factor(land2)) %>%
  right_join(combinations, by = c("set", "land1", "land2", "bullet1", "bullet2")) %>%
  filter(!(bullet1 == "Questioned" & bullet2 == "Known 1"),
         !(bullet1 == "Questioned" & bullet2 == "Known 2"),
         !(bullet1 == "Known 2" & bullet2 == "Known 1")) %>%
  arrange(rfscore) %>%
  mutate(case = factor(1:length(study))) %>%
  select(case, study, set, bullet1, land1, bullet2, land2, ccf:sum_peaks, samesource, rfscore)
```

@@

```{r}
# Save the test data as a .csv file
write.csv(hamby224_test, "../../../data/hamby224_test.csv", row.names = FALSE)
```

**Data for LIME Diagnostics Paper**

The code below takes the dataset `hamby224_test` and makes some formatting adjustments to prepare it to be used in the LIME diagnostics paper, so that it matches the training dataset published by Heike that was used to fit `rtrees`.

```{r}
# Change format of dataset to be used in the paper
paper_bullet_test_data <- 
  hamby224_test %>%
  mutate(study = str_remove(study, " "),
         set = str_remove(set, " "),
         bullet1 = str_replace(bullet1, "Known ", "B"),
         bullet2 = str_replace(bullet2, "Known ", "B"),
         bullet1 = str_replace(bullet1, "Questioned", "BUnk"),
         bullet2 = str_replace(bullet2, "Questioned", "BUnk"),
         land1 = str_replace(land1, "Land ", "L"),
         land2 = str_replace(land2, "Land ", "L")) %>%
  mutate(land_id1 = paste(study, set, bullet1, land1, sep = "-"),
         land_id2 = paste(study, set, bullet2, land2, sep = "-")) %>%
  select(-study, -set, -bullet1, -land1, -bullet2, -land2) %>%
  select(case, land_id1, land_id2, everything())

# ave the file to the folder where the LIME diagnostics paper is stored
write.csv(paper_bullet_test_data , "../../../diagnostics-paper/data-bullet-test.csv", row.names = FALSE)
```

@@

The random forest model that was fit to the Hamby training dataset is discussed in the paper "Automatic Matching of Bullet Land Impressions" by E. Hare Et. Al. The model is available through the GitHub version of the `bulletxtrctr` package. The link to the GitHub repository for `bulletxtrctr` is https://github.com/heike/bulletxtrctr.

The random forest model is stored in the package `bulletxtrctr` as a data object called `rtrees`. The paper does not include many details about the model fitting process, so I looked at some of the model features, which are printed below. These values told me that the model was fit using the predictive features of `ccf`, `rough_cor`, `D`, `sd_D`, `matches`, `mismatches`, `cms`, `non_cms`, and `sum_peaks` with 300 trees and the parameter of `mtry` set to 3, which is the "number of variables randomly sampled as candidates at each split" as explained in the `randomforest` package documentation.

@@

```{r}
# Create a dataframe with the random forest features ordered by importance
rf_features_ordered <- data.frame(feature = rownames(bulletxtrctr::rtrees$importance), 
                                  MeanDecreaseGini = bulletxtrctr::rtrees$importance) %>%
  arrange(desc(MeanDecreaseGini))
```
